---
title: "Pulling Data Sets"
author: "Hun Lee (sl4836), Jimmy Kelliher (jmk2303), Tanvir Khan (tk2886), Tucker Morgan (tlm2152), Zachary Katz (zak2132)"
date: "11/22/2021"
output: github_document
---

```{r setup, message = FALSE}
library(tidyverse)
library(httr)
library(gdata)
library(RCurl)
library(readxl)
```

The predictor variables in our analysis will come from United States Census data and the American Community Survey (ACS) 2019 five-year estimate via [IPUMS](https://usa.ipums.org/usa/). This data includes demographic and macroeconomic data by Public Use Microdata Area (PUMA) for New York City. Predictor variables include:

  * Rent 
  * Household income
  * Food stamps
  * High-speed internet
  * Family size
  * Sex
  * Age
  * Race / Ethnicity
  * Citizenship
  * Health coverage
  * Education
  * Employment Status
  * Personal income
  * Poverty status

Below, we bring in monthly health outcomes from [NYC Department of Health and Mental Hygiene](https://github.com/nychealth/coronavirus-data) (DOHMH) as of 11/18/2021. We will primarily analyze data related to COVID-19 hospitalizations and deaths.
```{r health outcomes import, message = FALSE}
hosp_rate_url = "https://raw.githubusercontent.com/nychealth/coronavirus-data/master/trends/hosprate-by-modzcta.csv"

hosp_rate_zcta <- read_csv(hosp_rate_url)

death_rate_url = "https://raw.githubusercontent.com/nychealth/coronavirus-data/master/trends/deathrate-by-modzcta.csv"

death_rate_zcta <- read_csv(death_rate_url)
```

```{r writing outcome data, eval = FALSE}
write_csv(hosp_rate_zcta, "./data/nyc_hosp_rate_zcta.csv")
write_csv(death_rate_zcta, "./data/nyc_death_rate_zcta.csv")
```

Below, we import vaccination data from [nyc.gov](https://www1.nyc.gov/site/doh/covid/covid-19-data-vaccines.page) containing information on the number and estimated percentage of NYC residents who are fully or partially vaccinated by Modified ZCTA (MODZCTA) as of 11/17/2021.
```{r importing nyc.gov vaccination data, message = FALSE}
vacc_url = "https://raw.githubusercontent.com/nychealth/covid-vaccine-data/main/people/coverage-by-modzcta-allages.csv"

nyc_vacc_zcta <- read_csv(vacc_url)
head(nyc_vacc_zcta)
```

```{r writing vaccination data to csv, eval = FALSE}
write_csv(nyc_vacc_zcta, "./data/nyc_vacc_zcta.csv")
```

Finally, we will import a crosswalk data set from [Baruch College](https://www.baruch.cuny.edu/confluence/display/geoportal/NYC+Geographies) to be utilized in relating ZCTAs and PUMAs.
```{r crosswalks for ZCTAs and PUMAs, message = FALSE}
zcta_puma_url = "http://faculty.baruch.cuny.edu/geoportal/resources/nyc_geog/nyc_zcta10_to_puma10.xls"

zcta_puma_cross <- read.xls(zcta_puma_url)
```

```{r writing crosswalks, eval = FALSE}
write_csv(zcta_puma_cross, "./data/zcta_puma_cross.csv")
```

