---
title: "exploratory_analysis_cleanin_TanvirK"
author: "Tanvir Khan" 
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# NOTES:
1. Broadband is now a binary variable with either the individual having internet or not having internet 

2. Birthplace is now a binary variable with either born in USA or not born in USA 

3. HISPAN is now a binary variable

2. Created a new column called yearly rent, did not delete monthly rent, need **GROUP INPUT** if we should have yearly_rent or monthly rent or both. 

# Before Recoding remove variables/columns that I feel is not needed for our project, 

# After removing the variables, now it's time to rename specific columns, I realized that some of these names for these variables are not user friendly:

```{r}
merged_data_clean <- 
  merged_data %>%
  select(-perwt, -multyear, -ancestr1, -ancestr2, -occ, -ind, -incwage, -incwelfr, -poverty, -occscore, -pwpuma00, -tranwork, -inctot) %>%
  
  rename(birthplace = bpl, 
         broadband_status = cihispeed, 
         foodstamp_status = foodstmp,
         education_status = educ,
         employment_status = empstat,
         family_size = famsize, 
         household_income = hhincome, 
         number_of_child = nchild, 
         laborforce_status = labforce,
         total_family_income = ftotinc, 
         years_in_USA = yrsusa1) %>%
  
  mutate(birthplace  = as.integer(birthplace)) %>%
  mutate(birthplace = case_when(
    birthplace %in% c(0:120) ~ "Born in USA",
    birthplace %in% c(150:950) ~ "Not Born in USA"
  )) %>% 
  mutate(broadband_status = 
           case_when(
             broadband_status %in% c("10", "11", "12", "13", "14", "15", "16", "17") ~ "Yes",
             broadband_status %in% c("20") ~ "No"
           )) %>% 
  mutate(race = recode(race, 
                    "1" = "White",
                    "2" = "Black/African America",
                    "3" = "American Indian or Alaska Native",
                    "4" = "Chinese",
                    "5" = "Japanese",
                    "6" = "Other Asian or Pacific Islander",
                    "7" = "Other race, nec",
                    "8" = "Two major races", 
                    "9" = "Three or more major races")) %>%
  mutate(race = as.factor(race)) %>% 
  mutate(sex = recode(sex,
                       "1" = "Male",
                       "2" = "Female")) %>%
  mutate(sex = as.factor(sex)) %>%
  mutate(hispan = case_when(
    hispan %in% c(1:4) ~ "YES",
    hispan %in% c(0) ~ "No",
    hispan %in% c(9) ~ "Not Reported")) %>%
  mutate(foodstamp_status  = recode(foodstamp_status ,
                          "0" = "N/A",
                          "1" = "No",
                          "2" = "Yes")) %>%
  mutate(foodstamp_status  = as.factor(foodstamp_status)) %>% 
  mutate(citizen = recode(citizen,
                          "0" = "N/A",
                          "1" = "Born abroad of American parents",
                          "2" = "Naturalized citizen",
                          "3" = "Not a citizen",
                          "4" = "Not a citizen, but has received first papers",
                          "5" = "Foreign born, citizenship status not reported",
                          "8" = "Illegible",
                          "9" = "Missing/blank")) %>%
  mutate(citizen = as.factor(citizen)) %>% 
  mutate(hcovany = recode(hcovany,
                               "1" = "No health insurance coverage",
                               "2" = "With health insurance coverage")) %>%
  mutate(hcovany = as.factor(hcovany)) %>% 
  mutate(hcovpriv = recode(hcovpriv,
                          "1" = "Without private health insurance coverage",
                          "2" = "With private health insurance coverage")) %>% 
  mutate(hcovpriv = as.factor(hcovpriv)) %>% 
  mutate(hcovpub = recode(hcovpub,
                          "1" = "Without public health insurance coverage",
                          "2" = "With public health insurance coverage")) %>% 
  mutate(hcovpub = as.factor(hcovpub)) %>% 
  mutate(education_status = recode(education_status,
                       "00" =	"N/A or no schooling",
                       "01" =	"Nursery school to grade 4",
                       "02" = "Grade 5, 6, 7, or 8",
                       "03" = "Grade 9",
                       "04" = "Grade 10",
                       "05" =	"Grade 11",
                       "06"	= "Grade 12",
                       "07" = "1 year of college",
                       "08" =	"2 years of college", 
                       "09" =	"3 years of college",
                       "10"	= "4 years of college", 
                       "11" =	"5+ years of college")) %>% 
  mutate(education_status = as.factor(education_status)) %>% 
  mutate(employment_status = recode(employment_status,
                          "0" = "N/A",
                          "1" = "Employed",
                          "2" = "Unemployed",
                          "3" = "Not in labor force")) %>%
  mutate(employment_status = as.factor(employment_status)) %>% 
  mutate(laborforce_status = recode(laborforce_status,
                           "0" = "N/A",
                           "1" = "No, not in the labor force",
                           "2" = "Yes, in the labor force")) %>%
  mutate(laborforce_status = as.factor(laborforce_status)) %>%
  mutate(yearly_rent = rent * 12)
```


