---
title: "Risk Scores and Clustering"
author: "Zachary Katz and Hun Lee"
date: "12/6/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries and defaults, echo = FALSE, message = FALSE, warning = FALSE}
# load necessary packages
library(tidyverse)
library(spatstat)
library(glmnet)
library(ggpubr)
library(GGally)
library(pvclust)
library(cluster)
library(factoextra)
library(tidymodels)
library(vip)
library(rsample)
library(themis)
library(parsnip)
library(workflows)
library(dials)

# set knitr defaults
knitr::opts_chunk$set(
    echo      = TRUE
  , message   = FALSE
  , fig.width = 6
  , fig.asp   = .6
  , out.width = "90%"
)

# set theme defaults
theme_set(
  theme_bw() +
  theme(
    legend.position = "bottom"
    , plot.title    = element_text(hjust = 0.5)
    , plot.subtitle = element_text(hjust = 0.5)    
    , plot.caption  = element_text(hjust = 0.0)
  )
)

# set color scale defaults
options(
    ggplot2.continuous.colour = "viridis"
  , ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```

```{r unzip, message = FALSE, warning = FALSE}
jimzip <- function(csv_file, path) {
  # create full path to csv file
  full_csv <- paste0(path, "/", csv_file)
  # append ".zip" to csv file
  zip_file <- paste0(full_csv, ".zip")
  # unzip file
  unzip(zip_file)
  # read csv
  data_extract <- read_csv(csv_file)
  # be sure to remove file once unzipped (it will live in working directory)
  on.exit(file.remove(csv_file))
  # output data
  data_extract
}

census_data <- jimzip("census_filtered.csv", "./data")
```

```{r merge, message = FALSE, warning = FALSE}
health_data <- read_csv("./data/outcome_puma.csv")

merged_data <- merge(census_data, health_data, by = "puma")
rm(census_data, health_data)
```

```{r clean, message = FALSE, warning = FALSE}
# Clean the merged census and outcomes data
cleaned_data = 
  merged_data %>% 
  # Remove variables less useful for analysis, including ones with high correlation with remaining variables
  select(-multyear, -ancestr1, -ancestr2, -labforce, -occ, -ind, -incwage, -occscore, -pwpuma00, -ftotinc, -hcovpub) %>% 
  # Remove duplicate rows
  distinct() %>% 
  # Rename variables
  rename(
    borough = countyfip,
    has_broadband = cihispeed,
    birthplace = bpl,
    education = educd,
    employment = empstat,
    personal_income = inctot,
    work_transport = tranwork,
    household_income = hhincome,
    on_foodstamps = foodstmp,
    family_size = famsize,
    num_children = nchild,
    US_citizen = citizen,
    puma_vacc_rate = puma_vacc_per,
    on_welfare = incwelfr,
    poverty_threshold = poverty
  ) %>% 
  # Recode variables according to data dictionary
  mutate(
    # Researched mapping for county
    borough = recode(
      borough,
      "5" = "Bronx",
      "47" = "Brooklyn",
      "61" = "Manhattan",
      "81" = "Queens",
      "85" = "Staten Island"
    ),
    rent = ifelse(
      rent == 9999, 0,
      rent
    ),
    household_income = ifelse(
      household_income %in% c(9999998,9999999), NA,
      household_income
    ),
    on_foodstamps = recode(
      on_foodstamps,
      "1" = "No",
      "2" = "Yes"
    ),
    has_broadband = case_when(
      has_broadband == "20" ~ "No",
      has_broadband != "20" ~ "Yes"
    ),
    sex = recode(
      sex,
      "1" = "Male",
      "2" = "Female"
    ),
    # Collapse Hispanic observation into race observation
    race = case_when(
      race == "1" ~ "White",
      race == "2" ~ "Black",
      race == "3" ~ "American Indian",
      race %in% c(4,5,6) ~ "Asian and Pacific Islander",
      race == 7 & hispan %in% c(1,2,3,4) ~ "Hispanic",
      race == 7 & hispan %in% c(0,9) ~ "Other",
      race %in% c(8,9) ~ "2+ races"
    ),
    birthplace = case_when(
      birthplace %in% 1:120 ~"US",
      birthplace %in% 121:950 ~ "Non-US",
      birthplace == 999 ~"Unknown"
    ),
    US_citizen = case_when(
      US_citizen %in% c(1,2) ~ "Yes",
      US_citizen %in% 3:8 ~"No",
      US_citizen %in% c(0,9) ~ "Unknown"
    ),
    # Chose languages based on highest frequency observed
    language = case_when(
      language == "1" ~ "English",
      language == "12" ~ "Spanish",
      language == "43" ~ "Chinese",
      language == "0" ~ "Unknown",
      language == "31" ~ "Hindi",
      !language %in% c(1,12,43,0,31) ~ "Other"
    ),
    # Collapse multiple health insurance variables into single variable
    health_insurance = case_when(
      hcovany == 1 ~ "None",
      hcovany == 2 & hcovpriv == 2 ~ "Private",
      hcovany == 2 & hcovpriv == 1 ~ "Public"
    ),
    education = case_when(
      education %in% 2:61 ~ "Less Than HS Graduate",
      education %in% 62:64 ~ "HS Graduate",
      education %in% 65:100 ~ "Some College",
      education %in% 110:113 ~ "Some College",
      education == 101 ~ "Bachelor's Degree",
      education %in% 114:116 ~ "Post-Graduate Degree",
      education %in% c(0,1,999) ~ "Unknown"
    ),
    employment = case_when(
      employment %in% c(0,3) ~ "Not in labor force",
      employment == 1 ~ "Employed",
      employment == 2 ~ "Unemployed"
    ),
    personal_income = ifelse(
      personal_income %in% c(9999998,9999999), NA,
      personal_income
    ),
    household_income = ifelse(
      household_income %in% c(9999998,9999999), NA,
      household_income
    ),
    on_welfare = case_when(
      on_welfare > 0 ~ "Yes",
      on_welfare == 0 ~ "No"
    ), 
    poverty_threshold = case_when(
      poverty_threshold >= 100 ~ "Above",
      poverty_threshold < 100 ~ "Below"
    ),
    work_transport = case_when(
      work_transport %in% c(31:37, 39) ~ "Public Transit",
      work_transport %in% c(10:20, 38) ~ "Private Vehicle",
      work_transport == 50 ~ "Bicycle",
      work_transport == 60 ~ "Walking",
      work_transport == 80 ~ "Worked From Home",
      work_transport %in% c(0, 70) ~ "Other"
    )
  ) %>% 
  # Convert hospitalization and death rates to pure percentages to match vax rate
  mutate(
    puma_hosp_rate = puma_hosp_rate / 1000,
    puma_death_rate = puma_death_rate / 1000
  ) %>% 
  # Eliminate columns no longer needed after transformation
  select(-hispan, -hcovany, -hcovpriv) %>% 
  # Relocate new columns
  relocate(health_insurance, .before = personal_income) %>% 
  relocate(poverty_threshold, .before = work_transport) %>% 
  relocate(on_welfare, .before = poverty_threshold) %>% 
  relocate(perwt, .before = cluster) %>% 
  # Create factor variables where applicable
  mutate(across(.cols = c(puma, borough, on_foodstamps, has_broadband, sex, race, birthplace, US_citizen, language, health_insurance, education, employment, on_welfare, poverty_threshold, work_transport), as.factor)) %>% 
  # Change levels of certain key factors for later analysis
  mutate(
    health_insurance = factor(health_insurance,
                              levels = c("None", "Public", "Private")),
    education = factor(education,
                       levels = c("Less Than HS Graduate", "HS Graduate", "Some College", "Bachelor's Degree", "Post-Graduate Degree", "Unknown"))
  )

rm(merged_data)
```

```{r puma summary, message = FALSE, warning = FALSE}
# Example data frame with weightings for summary stats over each PUMA
nyc_puma_summary = cleaned_data %>% 
  # Note: do we need to filter to one individual per household for household weightings?
  group_by(puma) %>%
  summarize(
    total_people = sum(perwt),
    median_household_income = weighted.median(household_income, hhwt, na.rm = TRUE),
    perc_foodstamps = sum(hhwt[on_foodstamps == "Yes"]) * 100 / sum(hhwt),
    perc_broadband = sum(hhwt[has_broadband == "Yes"]) * 100 / sum(hhwt),
    perc_male = sum(perwt[sex == "Male"]) * 100 / sum(perwt),
    median_age = weighted.median(age, perwt, na.rm = TRUE),
    perc_white = sum(perwt[race == "White"]) * 100 / sum(perwt),
    perc_foreign_born = sum(perwt[birthplace == "Non-US"]) * 100 / sum(perwt),
    perc_citizen = sum(perwt[US_citizen == "Yes"]) * 100 / sum(perwt),
    perc_english = sum(perwt[language == "English"]) * 100 / sum(perwt),
    perc_college = sum(perwt[education %in% c("Some College", "Bachelor's Degree", "Post-Graduate Degree")]) * 100 / sum(perwt),
    perc_unemployed = sum(perwt[employment == "Unemployed"]) * 100 / sum(perwt),
    perc_insured = sum(perwt[health_insurance %in% c("Private", "Public")]) * 100 / sum(perwt),
    median_personal_income = weighted.median(personal_income, perwt, na.rm = TRUE),
    perc_welfare = sum(perwt[on_welfare == "Yes"]) * 100 / sum(perwt),
    perc_poverty = sum(perwt[poverty_threshold == "Below"]) * 100 / sum(perwt),
    perc_public_transit = sum(perwt[work_transport == "Public Transit"]) * 100 / sum(perwt),
    covid_hosp_rate = median(puma_hosp_rate),
    covid_vax_rate = median(puma_vacc_rate),
    covid_death_rate = median(puma_death_rate)
  )
```

## Prediction Modeling

### Risk Scoring

We want to develop a method to score PUMAs on risk of not achieving herd immunity from vaccination. Let's say that herd immunity occurs at 70% vaccination rate, for our purposes.

```{r defining predictors and outcomes}
# 1 indicates BELOW 70% vaccination rate
logistic_df = nyc_puma_summary %>% 
  mutate(
    below_herd_vax = as.factor(ifelse(covid_vax_rate >= 70, 0, 1))
  ) %>% 
  select(-puma, -total_people, -covid_hosp_rate, -covid_death_rate, -covid_vax_rate)

# Define predictors matrix
x = model.matrix(below_herd_vax ~ ., logistic_df)[,-1]

# Define outcomes
y = logistic_df$below_herd_vax
```

```{r loading in new libraries}
library(caret) # Note: new libraries are loaded here due to potential function masking from the `caret` library
library(mlbench)
```


## Repeated Cross Validation
```{r finding optimal lambda}
set.seed(777)
vax_cv <- trainControl(method = "repeatedcv", number = 5, repeats = 100, 
                       savePredictions = T
                       )

# Goal is to find optimal lambda
lasso_model <- train(below_herd_vax ~ ., data = logistic_df,
                     method = "glmnet",
                     trControl = vax_cv,
                     tuneGrid = expand.grid(
                       .alpha = 1,
                       .lambda = seq(0.0001, 1, length = 100)),
                     family = "binomial")

lasso_model
```

## Result from training data
```{r resulting confusion matrix}
coef <- coef(lasso_model$finalModel, lasso_model$bestTune$lambda)

sub_lasso <-
  subset(lasso_model$pred, lasso_model$pred$lambda == lasso_model$bestTune$lambda)

caret::confusionMatrix(table(sub_lasso$pred, sub_lasso$obs))
```


The plot below shows coefficient estimates corresponding to a subset of the predictors in the data set.
```{r plotting predictors against lambda}
lambda  <- seq(0.0001, 1, length = 100)

lambda_opt = lasso_model$bestTune$lambda

result_plot <- broom::tidy(lasso_model$finalModel) %>% 
  select(term, lambda, estimate) %>% 
  complete(term, lambda, fill = list(estimate = 0) ) %>% 
  filter(term != "(Intercept)") %>% 
  ggplot(aes(x = log(lambda, 10), y = estimate, group = term, color = term)) + 
  geom_path() + 
  geom_vline(xintercept = log(lambda_opt, 10), color = "blue", size = 1.2) +
  theme(legend.position = "none") +
  labs(y = "Coefficient Estimate", title = "Coefficient Estimates for Varying Values of Lambda")
```

Below, we will obtain a risk prediction for each PUMA and visualize them.
```{r visualizing risk scores by PUMA, message = FALSE, warning = FALSE}
lambda <- lasso_model$bestTune$lambda
lasso_fit = glmnet(x, y, lambda = lambda, family = "binomial")
risk_predictions = (round((predict(lasso_fit, x, type = "response"))*100, 1))

puma <- nyc_puma_summary %>% 
  select(puma)

vax <- logistic_df %>% 
  select(below_herd_vax)

risk_prediction <- 
  bind_cols(puma, vax, as.vector(risk_predictions)) %>%
  rename(risk_prediciton = ...3)

risk_prediction %>%
  mutate(puma = fct_reorder(puma, risk_prediciton, .desc = TRUE)) %>%
  ggplot(aes(x = puma, y = risk_prediciton, fill = below_herd_vax)) + 
  geom_bar(stat  = "identity") + 
  geom_hline(yintercept = 50, linetype = "dashed", color = "red") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(title = "Predicting Risk Score of Vaccination Rate across PUMA",
       x = "PUMA", y = "Risk Score") + 
  scale_fill_manual(values = c("yellow", "purple"), labels = c("Below 70%", "Above 70%"))
```


## Clustering

We might try to see how our PUMAs cluster on predictors, and how those clusters are associated with distributions of particular outcomes. 

Let's start with K-means clustering.

```{r initial clustering}
# Define tibble of predictors only
predictors = nyc_puma_summary %>% 
  select(-puma, -total_people, -covid_hosp_rate, -covid_vax_rate, -covid_death_rate)

# Define tibble of outcomes only
outcomes = nyc_puma_summary %>% 
  select(covid_hosp_rate, covid_death_rate, covid_vax_rate)

# Define tibble of pumas only
pumas = nyc_puma_summary %>% 
  select(puma)

# Fit 3 clusters on predictors
kmeans_fit = 
  kmeans(x = predictors, centers = 3)

# Add clusters to data frame of predictors and bind with PUMA and outcomes data
predictors = 
  broom::augment(kmeans_fit, predictors)

# Bind columns
full_df = cbind(pumas, outcomes, predictors)

# Summary df
summary_df = full_df %>% 
  group_by(.cluster) %>% 
  summarize(
    median_hosp = median(covid_hosp_rate),
    median_death = median(covid_death_rate),
    median_vax = median(covid_vax_rate)
  )

# Plot predictor clusters against outcomes
# Example: try hospitalization vs vaccination
ggplot(data = full_df, aes(x = covid_hosp_rate, y = covid_vax_rate, color = .cluster)) + 
  geom_point() + 
  geom_point(data = summary_df, aes(x = median_hosp, y = median_vax), color = "black", size = 4) +
  geom_point(data = summary_df, aes(x = median_hosp, y = median_vax, color = .cluster), size = 2.75) +
  labs(x = "COVID Hospitalization Rate", y = "COVID Vaccination Rate", title = "COVID Hospitalization vs Vaccination Rates for each PUMA", color = "Cluster")
```

We could also scale predictors and omit NAs.

```{r scaled clustering}
# Scale predictors
for_clustering = predictors %>% 
  select(-.cluster) %>% 
  na.omit() %>% 
  scale()

# Evaluate Euclidean distances between observations
distance = get_dist(for_clustering)
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))

# Cluster with three centers
k_scaled = kmeans(for_clustering, centers = 3)

# Visualize cluster plot with reduction to two dimensions
fviz_cluster(k_scaled, data = for_clustering)

# Bind with outcomes and color clusters
full_df = for_clustering %>% 
  as_tibble() %>% 
  cbind(outcomes, pumas) %>% 
  mutate(
    cluster = k_scaled$cluster
  )

summary_df = full_df %>% 
  group_by(cluster) %>% 
  summarize(
    median_hosp = median(covid_hosp_rate),
    median_death = median(covid_death_rate),
    median_vax = median(covid_vax_rate)
  )

ggplot(data = full_df, aes(x = covid_hosp_rate, y = covid_vax_rate, color = factor(cluster))) + 
  geom_point() + 
  geom_point(data = summary_df, aes(x = median_hosp, y = median_vax), color = "black", size = 4) +
  geom_point(data = summary_df, aes(x = median_hosp, y = median_vax, color = factor(cluster)), size = 2.75) +
  labs(x = "COVID Hospitalization Rate", y = "COVID Vaccination Rate", title = "COVID Hospitalization vs Vaccination Rates for each PUMA", color = "Cluster")
```

We may want to evaluate this method's clustering quality as follows:

```{r}
# Check where elbow occurs using WSS method
fviz_nbclust(for_clustering, kmeans, method = "wss")

# Check for optimal number of clusters using silhouette method
fviz_nbclust(for_clustering, kmeans, method = "silhouette")

# Check number of clusters that minimize gap statistic
gap_stat = clusGap(for_clustering, FUN = kmeans, nstart = 25, K.max = 20, B = 50)
fviz_gap_stat(gap_stat)
```

It seems that one or two clusters may actually work better than three, when clustering on predictors. Below, we try it with two, just for fun.

```{r scaled - two clusters}
# Scale predictors
for_clustering = predictors %>% 
  select(-.cluster) %>% 
  na.omit() %>% 
  scale()

# Evaluate Euclidean distances between observations
distance = get_dist(for_clustering)
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))

# Cluster with two centers
k_scaled2 = kmeans(for_clustering, centers = 2)

# Visualize cluster plot with reduction to two dimensions
fviz_cluster(k_scaled2, data = for_clustering)

# Bind with outcomes and color clusters
full_df = for_clustering %>% 
  as_tibble() %>% 
  cbind(outcomes, pumas) %>% 
  mutate(
    cluster = k_scaled2$cluster
  )

summary_df = full_df %>% 
  group_by(cluster) %>% 
  summarize(
    median_hosp = median(covid_hosp_rate),
    median_death = median(covid_death_rate),
    median_vax = median(covid_vax_rate)
  )

ggplot(data = full_df, aes(x = covid_hosp_rate, y = covid_vax_rate, color = factor(cluster))) + 
  geom_point() + 
  geom_point(data = summary_df, aes(x = median_hosp, y = median_vax), color = "black", size = 4) +
  geom_point(data = summary_df, aes(x = median_hosp, y = median_vax, color = factor(cluster)), size = 2.75) +
  labs(x = "COVID Hospitalization Rate", y = "COVID Vaccination Rate", title = "COVID Hospitalization vs Vaccination Rates for each PUMA", color = "Cluster")
```