---
title: "P8105: Data Science I"
author: "Prediction Playground<br>Zachary Katz (UNI: zak2132)"
output:
  github_document:
    toc: TRUE
---

<!------------------------------------------------------------------------------
Preamble
------------------------------------------------------------------------------->

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# load necessary packages
library(tidyverse)
library(patchwork)
library(reshape2)
library(rstatix)
library(patchwork)
library(ggridges)
library(spatstat)
library(plotly)
library(ggpubr)
library(GGally)
library(glmnet)
library(modelr)
library(verification)

# set knitr defaults
knitr::opts_chunk$set(
    echo      = TRUE
  , message   = FALSE
  , fig.width = 6
  , fig.asp   = .6
  , out.width = "90%"
)

# set theme defaults
theme_set(
  theme_bw() +
  theme(
    legend.position = "bottom"
    , plot.title    = element_text(hjust = 0.5)
    , plot.subtitle = element_text(hjust = 0.5)    
    , plot.caption  = element_text(hjust = 0.0)
  )
)

# set color scale defaults
options(
    ggplot2.continuous.colour = "viridis"
  , ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```

```{r unzip, echo = FALSE, message = FALSE, warning = FALSE}
jimzip <- function(csv_file, path) {
  # create full path to csv file
  full_csv <- paste0(path, "/", csv_file)
  # append ".zip" to csv file
  zip_file <- paste0(full_csv, ".zip")
  # unzip file
  unzip(zip_file)
  # read csv
  data_extract <- read_csv(csv_file)
  # be sure to remove file once unzipped (it will live in working directory)
  on.exit(file.remove(csv_file))
  # output data
  data_extract
}

census_data <- jimzip("census_filtered.csv", "./data")
```

```{r merge, echo = FALSE, message = FALSE, warning = FALSE}
health_data <-
  read_csv("./data/outcome_puma.csv") %>%
  rename(puma = puma10)

merged_data <- merge(census_data, health_data, by = "puma")
rm(census_data)
```

```{r clean, echo = FALSE, message = FALSE, warning = FALSE}
# Clean the merged census and outcomes data
cleaned_data = 
  merged_data %>% 
  # Remove variables less useful for analysis, including ones with high correlation with remaining variables
  select(-multyear, -ancestr1, -ancestr2, -labforce, -occ, -ind, -incwage, -occscore, -pwpuma00, -ftotinc, -hcovpub) %>% 
  # Remove duplicate rows
  distinct() %>% 
  # Rename variables
  rename(
    borough = countyfip,
    has_broadband = cihispeed,
    birthplace = bpl,
    education = educd,
    employment = empstat,
    personal_income = inctot,
    work_transport = tranwork,
    household_income = hhincome,
    on_foodstamps = foodstmp,
    family_size = famsize,
    num_children = nchild,
    US_citizen = citizen,
    puma_vacc_rate = puma_vacc_per,
    on_welfare = incwelfr,
    poverty_threshold = poverty
  ) %>% 
  # Recode variables according to data dictionary
  mutate(
    # Researched mapping for county
    borough = recode(
      borough,
      "5" = "Bronx",
      "47" = "Brooklyn",
      "61" = "Manhattan",
      "81" = "Queens",
      "85" = "Staten Island"
    ),
    rent = ifelse(
      rent == 9999, 0,
      rent
    ),
    household_income = ifelse(
      household_income %in% c(9999998,9999999), NA,
      household_income
    ),
    on_foodstamps = recode(
      on_foodstamps,
      "1" = "No",
      "2" = "Yes"
    ),
    has_broadband = case_when(
      has_broadband == "20" ~ "No",
      has_broadband != "20" ~ "Yes"
    ),
    sex = recode(
      sex,
      "1" = "Male",
      "2" = "Female"
    ),
    # Collapse Hispanic observation into race observation
    race = case_when(
      race == "1" ~ "White",
      race == "2" ~ "Black",
      race == "3" ~ "American Indian",
      race %in% c(4,5,6) ~ "Asian and Pacific Islander",
      race == 7 & hispan %in% c(1,2,3,4) ~ "Hispanic",
      race == 7 & hispan %in% c(0,9) ~ "Other",
      race %in% c(8,9) ~ "2+ races"
    ),
    birthplace = case_when(
      birthplace %in% 1:120 ~"US",
      birthplace %in% 121:950 ~ "Non-US",
      birthplace == 999 ~"Unknown"
    ),
    US_citizen = case_when(
      US_citizen %in% c(1,2) ~ "Yes",
      US_citizen %in% 3:8 ~"No",
      US_citizen %in% c(0,9) ~ "Unknown"
    ),
    # Chose languages based on highest frequency observed
    language = case_when(
      language == "1" ~ "English",
      language == "12" ~ "Spanish",
      language == "43" ~ "Chinese",
      language == "0" ~ "Unknown",
      language == "31" ~ "Hindi",
      !language %in% c(1,12,43,0,31) ~ "Other"
    ),
    # Collapse multiple health insurance variables into single variable
    health_insurance = case_when(
      hcovany == 1 ~ "None",
      hcovany == 2 & hcovpriv == 2 ~ "Private",
      hcovany == 2 & hcovpriv == 1 ~ "Public"
    ),
    education = case_when(
      education %in% 2:61 ~ "Less Than HS Graduate",
      education %in% 62:64 ~ "HS Graduate",
      education %in% 65:100 ~ "Some College",
      education %in% 110:113 ~ "Some College",
      education == 101 ~ "Bachelor's Degree",
      education %in% 114:116 ~ "Post-Graduate Degree",
      education %in% c(0,1,999) ~ "Unknown"
    ),
    employment = case_when(
      employment %in% c(0,3) ~ "Not in labor force",
      employment == 1 ~ "Employed",
      employment == 2 ~ "Unemployed"
    ),
    personal_income = ifelse(
      personal_income %in% c(9999998,9999999), NA,
      personal_income
    ),
    household_income = ifelse(
      household_income %in% c(9999998,9999999), NA,
      household_income
    ),
    on_welfare = case_when(
      on_welfare > 0 ~ "Yes",
      on_welfare == 0 ~ "No"
    ), 
    poverty_threshold = case_when(
      poverty_threshold >= 100 ~ "Above",
      poverty_threshold < 100 ~ "Below"
    ),
    work_transport = case_when(
      work_transport %in% c(31:37, 39) ~ "Public Transit",
      work_transport %in% c(10:20, 38) ~ "Private Vehicle",
      work_transport == 50 ~ "Bicycle",
      work_transport == 60 ~ "Walking",
      work_transport == 80 ~ "Worked From Home",
      work_transport %in% c(0, 70) ~ "Other"
    )
  ) %>% 
  # Convert hospitalization and death rates to pure percentages to match vax rate
  mutate(
    puma_hosp_rate = puma_hosp_rate / 1000,
    puma_death_rate = puma_death_rate / 1000
  ) %>% 
  # Eliminate columns no longer needed after transformation
  select(-hispan, -hcovany, -hcovpriv) %>% 
  # Relocate new columns
  relocate(health_insurance, .before = personal_income) %>% 
  relocate(poverty_threshold, .before = work_transport) %>% 
  relocate(on_welfare, .before = poverty_threshold) %>% 
  relocate(perwt, .before = cluster) %>% 
  # Create factor variables where applicable
  mutate(across(.cols = c(puma, borough, on_foodstamps, has_broadband, sex, race, birthplace, US_citizen, language, health_insurance, education, employment, on_welfare, poverty_threshold, work_transport), as.factor)) %>% 
  # Change levels of certain key factors for later analysis
  mutate(
    health_insurance = factor(health_insurance,
                              levels = c("None", "Public", "Private")),
    education = factor(education,
                       levels = c("Less Than HS Graduate", "HS Graduate", "Some College", "Bachelor's Degree", "Post-Graduate Degree", "Unknown"))
  )

```

```{r}
# Example data frame with weightings for summary stats over each PUMA
nyc_puma_summary = cleaned_data %>% 
  # Note: do we need to filter to one individual per household for household weightings?
  group_by(puma) %>%
  summarize(
    total_people = sum(perwt),
    median_household_income = weighted.median(household_income, hhwt, na.rm = TRUE),
    perc_foodstamps = sum(hhwt[on_foodstamps == "Yes"]) * 100 / sum(hhwt),
    perc_broadband = sum(hhwt[has_broadband == "Yes"]) * 100 / sum(hhwt),
    perc_male = sum(perwt[sex == "Male"]) * 100 / sum(perwt),
    median_age = weighted.median(age, perwt, na.rm = TRUE),
    perc_white = sum(perwt[race == "White"]) * 100 / sum(perwt),
    perc_foreign_born = sum(perwt[birthplace == "Non-US"]) * 100 / sum(perwt),
    perc_citizen = sum(perwt[US_citizen == "Yes"]) * 100 / sum(perwt),
    perc_english = sum(perwt[language == "English"]) * 100 / sum(perwt),
    perc_college = sum(perwt[education %in% c("Some College", "Bachelor's Degree", "Post-Graduate Degree")]) * 100 / sum(perwt),
    perc_unemployed = sum(perwt[employment == "Unemployed"]) * 100 / sum(perwt),
    perc_insured = sum(perwt[health_insurance %in% c("Private", "Public")]) * 100 / sum(perwt),
    median_personal_income = weighted.median(personal_income, perwt, na.rm = TRUE),
    perc_welfare = sum(perwt[on_welfare == "Yes"]) * 100 / sum(perwt),
    perc_poverty = sum(perwt[poverty_threshold == "Below"]) * 100 / sum(perwt),
    perc_public_transit = sum(perwt[work_transport == "Public Transit"]) * 100 / sum(perwt),
    covid_hosp_rate = median(puma_hosp_rate),
    covid_vax_rate = median(puma_vacc_rate),
    covid_death_rate = median(puma_death_rate)
  )
```


## Prediction Modeling

### Risk scoring

We want to develop a method to score PUMAs on risk of not achieving herd immunity from vaccination. Let's say that herd immunity occurs at 70% vaccination rate, for our purposes.

```{r}
# Define binary outcome and set of predictors
# 1 indicates BELOW 70% vaccination rate
logistic_df = nyc_puma_summary %>% 
  mutate(
    below_herd_vax = ifelse(covid_vax_rate >= 70, 0, 1)
  ) %>% 
  select(-puma, -total_people, -covid_hosp_rate, -covid_death_rate, -covid_vax_rate)

# Define predictors and outcome
x = model.matrix(below_herd_vax ~ ., logistic_df)[,-1]
y = logistic_df$below_herd_vax
```

Then, we want to develop the penalized (logistic) regression model.

```{r}
# Define a grid of possible tuning parameters
lambda = 10 ^ seq(3, -2, -0.1)

# Build model
lasso_fit = cv.glmnet(x, y, lambda = lambda, family = "binomial")

# Define minimal lambda
lambda_opt = lasso_fit$lambda.min
```

Let's check out the coefficients from the optimal model:

```{r}
# Calculate model with optimal tuning parameter
lasso_optimal = glmnet(x, y, lambda = lambda_opt, family = "binomial")

# Determine coefficients
lasso_optimal %>% broom::tidy()
```

Now, let's see the predictions our model makes for each PUMA and assess fit.

```{r}
# Model predictions for each PUMA
predictions = predict(lasso_optimal, x, type = "class")

# Confusion matrix
confusionMatrix(data = factor(predictions), reference = factor(y))
```

For raw risk scores, we can use the fitted probabilities for our binomial classifier.

```{r}
# Fitted probabilities for binomial classifier, transformed into percentages
risk_predictions = round((predict(lasso_optimal, x, type = "response"))*100, 1)

# List of PUMAS
pumas = nyc_puma_summary %>% 
  select(puma)

# Join with PUMA and create table
puma_risk_score = cbind(pumas, risk_predictions) %>% 
  knitr::kable()

# In-sample performance of model using ROC plot
roc.plot(y == 1, risk_predictions/100)$roc.vol
```

## PLAYGROUND: PLEASE IGNORE FOR NOW

### K-Means Clustering

Can we cluster our PUMAs according to predictors, and see how well the clustering predicts outcomes?

```{r}
# K-Means Clustering using predictors

predictors_only = nyc_puma_summary %>% 
  select(-puma, -total_people, -covid_hosp_rate, -covid_vax_rate, -covid_death_rate)

outcomes_only_and_key = nyc_puma_summary %>% 
  select(median_household_income, puma, covid_hosp_rate, covid_death_rate, covid_vax_rate)

kmeans_fit = 
  kmeans(x = predictors_only, centers = 3)

predictors_only = 
  broom::augment(kmeans_fit, predictors_only) 

puma_clusters = predictors_only %>% 
  left_join(outcomes_only_and_key, by = "median_household_income") %>% 
  select(puma, covid_hosp_rate.x, covid_death_rate.x, covid_vax_rate.x, covid_death_rate.x, .cluster)

puma_cluster_summary = puma_clusters %>% 
  group_by(.cluster) %>% 
  summarize(
    median_hosp = median(covid_hosp_rate.x),
    median_death = median(covid_death_rate.x),
    median_vax = median(covid_vax_rate.x)
  )

ggplot(data = puma_cluster_summary, aes(x = median_hosp, y = median_death, color = median_vax)) + 
  geom_point()

ggplot(data = puma_clusters, aes(x = covid_hosp_rate.x, y=covid_vax_rate.x, color = .cluster)) + 
  geom_point() + 
  geom_point(aes(x = 1.06, y = 58.4), color = "red", size = 4) + 
  geom_point(aes(x = 1.06, y = 49.4), color = "red", size = 4) + 
  geom_point(aes(x = 0.493, y = 73.5), color = "red", size = 4)
```

How about with four clusters?

```{r}
# K-Means Clustering

predictors_only = nyc_puma_summary %>% 
  select(-puma, -total_people, covid_hosp_rate, covid_vax_rate, covid_death_rate)

outcomes_only_and_key = nyc_puma_summary %>% 
  select(median_household_income, puma, covid_hosp_rate, covid_death_rate, covid_vax_rate)

kmeans_fit = 
  kmeans(x = predictors_only, centers = 4)

predictors_only = 
  broom::augment(kmeans_fit, predictors_only) 

puma_clusters = predictors_only %>% 
  left_join(outcomes_only_and_key, by = "median_household_income") %>% 
  select(puma, covid_hosp_rate.x, covid_death_rate.x, covid_vax_rate.x, covid_death_rate.x, .cluster)

puma_cluster_summary = puma_clusters %>% 
  group_by(.cluster) %>% 
  summarize(
    median_hosp = median(covid_hosp_rate.x),
    median_death = median(covid_death_rate.x),
    median_vax = median(covid_vax_rate.x)
  )

ggplot(data = puma_cluster_summary, aes(x = median_hosp, y = median_death, color = median_vax)) + 
  geom_point()

ggplot(data = puma_clusters, aes(x = covid_hosp_rate.x, y=covid_vax_rate.x, color = .cluster)) + 
  geom_point() + 
  geom_point(aes(x = 0.961, y = 54.3), color = "red", size = 4) + 
  geom_point(aes(x = 0.493, y = 73.5), color = "red", size = 4) + 
  geom_point(aes(x = 1.13, y = 59.6), color = "red", size = 4) + 
  geom_point(aes(x = 1.13, y = 50.2), color = "red", size = 4)
```

This is based only on using predictors to cluster the data. What if we clustered based solely on hospitalization, death, and vaccination rate? Let's try with three clusters:

```{r}
# K-Means Clustering on outcomes only
# 3 clusters only

outcomes_only = nyc_puma_summary %>% 
  select(puma, covid_hosp_rate, covid_death_rate, covid_vax_rate)

outcomes_without_puma = outcomes_only %>% 
  select(-puma)

kmeans_fit = 
  kmeans(x = outcomes_without_puma, centers = 3)

outcomes_without_puma = 
  broom::augment(kmeans_fit, outcomes_without_puma)

outcomes_without_puma %>% 
  ggplot(aes(x = covid_hosp_rate, y = covid_vax_rate, color = .cluster)) + 
  geom_point()
```

Then try with four clusters:

```{r}
# K-Means Clustering on outcomes only
# 4 clusters only

outcomes_only = nyc_puma_summary %>% 
  select(puma, covid_hosp_rate, covid_death_rate, covid_vax_rate)

outcomes_without_puma = outcomes_only %>% 
  select(-puma)

kmeans_fit = 
  kmeans(x = outcomes_without_puma, centers = 4)

outcomes_without_puma = 
  broom::augment(kmeans_fit, outcomes_without_puma)

outcomes_without_puma %>% 
  ggplot(aes(x = covid_hosp_rate, y = covid_vax_rate, color = .cluster)) + 
  geom_point()
```

(Not really better than three clusters.)

Finally, let's try clustering on both predictors *and* on outcomes (start with 4 clusters).

```{r}
cluster_df = nyc_puma_summary %>% 
  select(-puma, -total_people)

kmeans_fit = 
  kmeans(x = cluster_df, centers = 4)

cluster_df = 
  broom::augment(kmeans_fit, cluster_df)

cluster_df %>% 
  ggplot(aes(x = covid_hosp_rate, y = covid_vax_rate, color = .cluster)) + 
  geom_point()
```

### Hierarchical clustering

```{r}
library(pvclust)

predictors_only = predictors_only %>% 
  select(-covid_hosp_rate, -covid_vax_rate, -covid_death_rate, -.cluster)

fit = pvclust(predictors_only, method.hclust = "ward", method.dist = "euclidean")
plot(fit)
pvrect(fit, alpha = 0.95)
```

```{r}
library(tidyverse)
library(cluster)
library(factoextra)

for_clustering = nyc_puma_summary %>% 
  na.omit() %>% 
  select(-puma) %>% 
  scale()

distance = get_dist(for_clustering)
fviz_dist(distance, gradient = list(low = "#00AFBB", mid = "white", high = "#FC4E07"))

k2 = kmeans(for_clustering, centers = 3, nstart = 25)

fviz_cluster(k2, data = for_clustering)

for_clustering %>% 
  as_tibble() %>% 
  mutate(
    cluster = k2$cluster
  ) %>% 
  ggplot(aes(covid_hosp_rate, covid_vax_rate, color = factor(cluster))) + 
  geom_point()
```

```{r}
set.seed(123)

fviz_nbclust(for_clustering, kmeans, method = "wss")
fviz_nbclust(for_clustering, kmeans, method = "silhouette")
gap_stat = clusGap(for_clustering, FUN = kmeans, nstart = 25, K.max = 10, B = 50)
fviz_gap_stat(gap_stat)
```