---
title: "exploratory_analysis_TK"
output: github_document
---

```{r}
library(tidyverse)
library(patchwork)
library(ggridges)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!------------------------------------------------------------------------------
Preamble
------------------------------------------------------------------------------->

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# load necessary packages
library(tidyverse)
# set knitr defaults
knitr::opts_chunk$set(
    echo      = TRUE
  , message   = FALSE
  , fig.width = 6
  , fig.asp   = .6
  , out.width = "90%"
)
# set theme defaults
theme_set(
  theme_bw() +
  theme(
    legend.position = "bottom"
    , plot.title    = element_text(hjust = 0.5)
    , plot.subtitle = element_text(hjust = 0.5)    
    , plot.caption  = element_text(hjust = 0.0)
  )
)
# set color scale defaults
options(
    ggplot2.continuous.colour = "viridis"
  , ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```

<!------------------------------------------------------------------------------
Overview
------------------------------------------------------------------------------->

# Unzipping the Census Data

```{r}
jimzip <- function(csv_file, path) {
  # create full path to csv file
  full_csv <- paste0(path, "/", csv_file)
  # append ".zip" to csv file
  zip_file <- paste0(full_csv, ".zip")
  # unzip file
  unzip(zip_file)
  # read csv
  data_extract <- read_csv(csv_file)
  # be sure to remove file once unzipped (it will live in working directory)
  on.exit(file.remove(csv_file))
  # output data
  data_extract
}
census_data <- jimzip("census_filtered.csv", "./data")
head(census_data) %>%  knitr::kable()
```

# Merging the Outcome Data

```{r}
health_data <-
  read_csv("./data/outcome_puma.csv") %>%
  rename(puma = puma10)
merged_data <- merge(census_data, health_data, by = "puma")
rm(census_data)
head(merged_data) %>% knitr::kable()
```

## Cleaning the Data

```{r}
# Clean the merged census and outcomes data
cleaned_data = 
  merged_data %>% 
  # Remove variables less useful for analysis, including ones with high correlation with remaining variables
  select(-multyear, -ancestr1, -ancestr2, -labforce, -occ, -ind, -incwage, -occscore, -pwpuma00, -ftotinc, -hcovpub) %>% 
  # Remove duplicate rows
  distinct() %>% 
  # Rename variables
  rename(
    borough = countyfip,
    has_broadband = cihispeed,
    birthplace = bpl,
    education = educd,
    employment = empstat,
    personal_income = inctot,
    work_transport = tranwork,
    household_income = hhincome,
    on_foodstamps = foodstmp,
    family_size = famsize,
    num_children = nchild,
    US_citizen = citizen,
    puma_vacc_rate = puma_vacc_per,
    on_welfare = incwelfr,
    poverty_threshold = poverty
  ) %>% 
  # Recode variables according to data dictionary
  mutate(
    # Researched mapping for county
    borough = recode(
      borough,
      "5" = "Bronx",
      "47" = "Brooklyn",
      "61" = "Manhattan",
      "81" = "Queens",
      "85" = "Staten Island"
    ),
    rent = ifelse(
      rent == 9999, 0,
      rent
    ),
    household_income = ifelse(
      household_income %in% c(9999998,9999999), NA,
      household_income
    ),
    on_foodstamps = recode(
      on_foodstamps,
      "1" = "No",
      "2" = "Yes"
    ),
    has_broadband = case_when(
      has_broadband == "20" ~ "No",
      has_broadband != "20" ~ "Yes"
    ),
    sex = recode(
      sex,
      "1" = "Male",
      "2" = "Female"
    ),
    # Collapse Hispanic observation into race observation
    race = case_when(
      race == "1" ~ "White",
      race == "2" ~ "Black",
      race == "3" ~ "American Indian",
      race %in% c(4,5,6) ~ "Asian and Pacific Islander",
      race == 7 & hispan %in% c(1,2,3,4) ~ "Hispanic",
      race == 7 & hispan %in% c(0,9) ~ "Other",
      race %in% c(8,9) ~ "2+ races"
    ),
    birthplace = case_when(
      birthplace %in% 1:120 ~"US",
      birthplace %in% 121:950 ~ "Non-US",
      birthplace == 999 ~"Unknown"
    ),
    US_citizen = case_when(
      US_citizen %in% c(1,2) ~ "Yes",
      US_citizen %in% 3:8 ~"No",
      US_citizen %in% c(0,9) ~ "Unknown"
    ),
    # Chose languages based on highest frequency observed
    language = case_when(
      language == "1" ~ "English",
      language == "12" ~ "Spanish",
      language == "43" ~ "Chinese",
      language == "0" ~ "Unknown",
      language == "31" ~ "Hindi",
      !language %in% c(1,12,43,0,31) ~ "Other"
    ),
    # Collapse multiple health insurance variables into single variable
    health_insurance = case_when(
      hcovany == 1 ~ "None",
      hcovany == 2 && hcovpriv == 2 ~ "Private",
      hcovany == 2 && hcovpriv == 1 ~ "Public"
    ),
    education = case_when(
      education %in% 2:61 ~ "Less Than HS Graduate",
      education %in% 62:64 ~ "HS Graduate",
      education %in% 65:100 ~ "Some College",
      education %in% 110:113 ~ "Some College",
      education == 101 ~ "Bachelor's Degree",
      education %in% 114:116 ~ "Post-Graduate Degree",
      education %in% c(0,1,999) ~ "Unknown"
    ),
    employment = case_when(
      employment %in% c(0,3) ~ "Not in labor force",
      employment == 1 ~ "Employed",
      employment == 2 ~ "Unemployed"
    ),
    personal_income = ifelse(
      personal_income %in% c(9999998,9999999), NA,
      personal_income
    ),
    household_income = ifelse(
      household_income %in% c(9999998,9999999), NA,
      household_income
    ),
    on_welfare = case_when(
      on_welfare > 0 ~ "Yes",
      on_welfare == 0 ~ "No"
    ), 
    poverty_threshold = case_when(
      poverty_threshold >= 100 ~ "Above",
      poverty_threshold < 100 ~ "Below"
    ),
    work_transport = case_when(
      work_transport %in% c(31:37, 39) ~ "Public Transit",
      work_transport %in% c(10:20, 38) ~ "Private Vehicle",
      work_transport == 50 ~ "Bicycle",
      work_transport == 60 ~ "Walking",
      work_transport == 80 ~ "Worked From Home",
      work_transport %in% c(0, 70) ~ "Other"
    )
  ) %>% 
  # Convert hospitalization and death rates to pure percentages to match vax rate
  mutate(
    puma_hosp_rate = puma_hosp_rate, 
    puma_death_rate = puma_death_rate / 1000
  ) %>% 
  # Eliminate columns no longer needed after transformation
  select(-hispan, -hcovany, -hcovpriv) %>% 
  # Relocate new columns
  relocate(health_insurance, .before = personal_income) %>% 
  relocate(poverty_threshold, .before = work_transport) %>% 
  relocate(on_welfare, .before = poverty_threshold) %>% 
  relocate(perwt, .before = cluster) %>% 
  # Create factor variables where applicable
  mutate(across(.cols = c(puma, borough, on_foodstamps, has_broadband, sex, race, birthplace, US_citizen, language, health_insurance, education, employment, on_welfare, poverty_threshold, work_transport), as.factor))
```

The Census also doesn't reach everyone; it samples the population. As a result, we need to weight our statistics by how many households look like a particular household observed, or how many persons look like a particular person observed. We also want to group by PUMA.

```{r}
# Check total people in NYC by summing over person-weight column
sum(cleaned_data$perwt)
# Load spatstat library
library(spatstat)
# Example data frame with weightings for summary stats over each PUMA
nyc_puma_summary = cleaned_data %>% 
  # Note: do we need to filter to one individual per household for household weightings?
  group_by(puma) %>%
  summarize(
    median_household_income = weighted.median(household_income, hhwt, na.rm = TRUE),
    perc_foodstamps = sum(hhwt[on_foodstamps == "Yes"]) * 100 / sum(hhwt),
    perc_broadband = sum(hhwt[has_broadband == "Yes"]) * 100 / sum(hhwt),
    perc_male = sum(perwt[sex == "Male"]) * 100 / sum(perwt),
    median_age = weighted.median(age, perwt, na.rm = TRUE),
    perc_white = sum(perwt[race == "White"]) * 100 / sum(perwt),
    perc_non_white = sum(perwt[race != "White"]) * 100 / sum(perwt), 
    perc_foreign_born = sum(perwt[birthplace == "Non-US"]) * 100 / sum(perwt),
    perc_us_born = sum(perwt[birthplace == "US"]) * 100 / sum(perwt),
    perc_citizen = sum(perwt[US_citizen == "Yes"]) * 100 / sum(perwt),
    perc_english = sum(perwt[language == "English"]) * 100 / sum(perwt),
    perc_non_english =  sum(perwt[language != "English"]) * 100 / sum(perwt), 
    perc_college = sum(perwt[education %in% c("Some College", "Bachelor's Degree", "Post-Graduate Degree")]) * 100 / sum(perwt),
    perc_not_college = sum(perwt[education %in% c("Less Than HS Graduate", "HS Graduate")]) * 100 / sum(perwt),
    perc_unemployed = sum(perwt[employment == "Unemployed"]) * 100 / sum(perwt),
    perc_insured = sum(perwt[health_insurance %in% c("Private", "Public")]) * 100 / sum(perwt),
    median_personal_income = weighted.median(personal_income, perwt, na.rm = TRUE),
    perc_welfare = sum(perwt[on_welfare == "Yes"]) * 100 / sum(perwt),
    perc_poverty = sum(perwt[poverty_threshold == "Below"]) * 100 / sum(perwt),
    perc_public_transit = sum(perwt[work_transport == "Public Transit"]) * 100 / sum(perwt),
    covid_hosp_rate = median(puma_hosp_rate),
    covid_vax_rate = median(puma_vacc_rate),
    covid_death_rate = median(puma_death_rate)
  )
# View summary data across PUMAs
nyc_puma_summary %>% 
  View()
```


**Exploratory Analysis (1st Attempt):**

*I realized that in nyc_puma_summary there is no borough name, so that may affect when it comes to the graph output*

```{r}
borough_info <- 
  nyc_puma_summary %>% 
  mutate(
    Borough = (
      case_when(
        str_detect(puma, "37") ~ "Bronx",
        str_detect(puma, "38") ~ "Manhattan",
        str_detect(puma, "39") ~ "Staten Island",
        str_detect(puma, "40") ~ "Brooklyn",
        str_detect(puma, "41") ~ "Queens")))
```


**SCATTERPLOTS**

COVID-Death Rate Across Puma (smallest to largest)
```{r}
borough_death_rate <- 
  borough_info %>%
  ggplot(aes(x = fct_reorder(puma, covid_death_rate), 
             y = covid_death_rate)) + 
  geom_point(aes(color = Borough)) + 
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, hjust = 0.5)) + 
  labs(
    title = "COVID Death Rates Across Puma", 
    subtitle = "Ordered: smallest to largest rate", 
    x = "Puma",
    y = "COVID death rate"
  ) +
  viridis::scale_color_viridis(
    name = "Boroughs in NYC", 
    discrete = TRUE
  ) +
  theme(legend.position = "right")
```




Covid-Vaccination Rate Across Puma (smallest to largest)
```{r}
borough_vaccination_rate <-
  borough_info %>%
  ggplot(aes(x = fct_reorder(puma, covid_vax_rate), 
             y = covid_vax_rate)) + 
  geom_point(aes(color = Borough)) + 
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, hjust = 0.5)) + 
  labs(
    title = "COVID Vaccination Rates Across Puma", 
    subtitle = "Ordered: smallest to largest rate", 
    x = "Puma",
    y = "COVID Vaccination rate"
  ) +
  viridis::scale_color_viridis(
    name = "Boroughs in NYC", 
    discrete = TRUE
  ) +
  theme(legend.position = "right")
``` 



COVID-Hospitalization Rate Across Puma (smallest to largest)
```{r}
borough_hospitalization_rate <-
  borough_info %>%
  ggplot(aes(x = fct_reorder(puma, covid_hosp_rate), 
             y = covid_hosp_rate)) + 
  geom_point(aes(color = Borough)) + 
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, hjust = 0.5)) + 
  labs(
    title = "COVID Hospitalization Rates Across Puma", 
    subtitle = "Ordered: smallest to largest rate", 
    x = "Puma",
    y = "COVID Hospitalization rate per 100,000"
  ) +
  viridis::scale_color_viridis(
    name = "Boroughs in NYC", 
    discrete = TRUE
  ) +
  theme(legend.position = "right")
```




*BOXPLOT - Distribution of the data for each borough (hospitalization, death, vaccination)*


Distribution of death rate for each borough 
```{r}
borough_death_dist <-
  borough_info %>% 
  ggplot(aes(x = reorder(Borough, covid_death_rate),
             y = covid_death_rate)) + 
  geom_boxplot(aes(color = Borough)) + 
  labs(
    title = "COVID Death Rate for Each Burrough", 
    x = "NYC Borough",
    y = "COVID Death Rate"
  ) + 
  viridis::scale_color_viridis(
    name = "Boroughs in NYC", 
    discrete = TRUE
  ) +
  theme(legend.position = "right")
```



Distribution of vaccination rate for each borough 
```{r}
borough_vax_dist <-
  borough_info %>% 
  ggplot(aes(x = reorder(Borough, covid_vax_rate),
             y = covid_vax_rate)) + 
  geom_boxplot(aes(color = Borough)) + 
  labs(
    title = "COVID Vaccination Rate for Each Burrough", 
    x = "NYC Borough",
    y = "COVID Vaccination Rate"
  ) + 
  viridis::scale_color_viridis(
    name = "Boroughs in NYC", 
    discrete = TRUE
  ) +
  theme(legend.position = "right")
```



Distribution of hospitalization rate for each borough 
```{r}
borough_hosp_dist <-
  borough_info %>% 
  ggplot(aes(x = reorder(Borough, covid_hosp_rate), 
             y = covid_hosp_rate)) + 
  geom_boxplot(aes(color = Borough)) + 
  labs(
    title = "COVID Hospitalization Rate for Each Burrough", 
    x = "NYC Borough",
    y = "COVID Hospitalization Rate"
  ) + 
  viridis::scale_color_viridis(
    name = "Boroughs in NYC", 
    discrete = TRUE
  ) +
  theme(legend.position = "right")
```






*Socioeconomic information compared with Vaccination Rate* 

Income and Vaccination Rate 
```{r}
income_vax_relation <-
  borough_info %>% 
  ggplot(aes(x = median_household_income,
             y = covid_vax_rate)) + 
  geom_point() + 
  geom_smooth(
    method = "loess", 
    color = 'red',
    se = FALSE) +
  labs(
    title = "Median Household-Income VS.COVID Vaccination Rate",
    x = "Median household Income", 
    y = "Covid Vaccination Rate"
  ) +
  scale_x_continuous(
    breaks = c(50000, 75000, 100000, 125000, 150000, 175000, 200000),
    labels = c("$50,000", "$75,000", "$100,000", "$125,000", "$150,000", "$175,000", "$200,000")) 
```








College and Vaccination Rate 
```{r}
college_vax_rel <- 
  borough_info %>% 
  ggplot(aes(x = perc_college,
             y = covid_vax_rate)) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE) + 
  labs(
    title = "College and Vaccination",
    x = "College", 
    y = "Covid Vaccination Rate"
  ) +
  theme(legend.position = "none")
  
```


Non-College and Vaccination Rate 
```{r}
non_college_vax_rel <-
  borough_info %>% 
  ggplot(aes(x = perc_not_college,
             y = covid_vax_rate)) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE) +
  labs(
    title = "Non-College and Vaccination",
    x = "Non-College", 
    y = "Covid Vaccination Rate"
  ) + 
  theme(legend.position = "right")
```


Compare the plots side by side 
```{r}
college_vax_rel + non_college_vax_rel
```


College and Hospitalization Rate 
```{r}
college_hosp_rel <- 
  borough_info %>% 
  ggplot(aes(x = perc_college,
             y = covid_hosp_rate)) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE) + 
  labs(
    title = "College and Hospitalization",
    x = "College", 
    y = "Covid Hospitalization Rate per 100,000"
  ) +
  theme(legend.position = "right")
```



Non-College Hospitalization Rate 
```{r}
noncollege_hosp_rel <- 
  borough_info %>% 
  ggplot(aes(x = perc_not_college,
             y =  covid_hosp_rate)) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE) + 
  labs(
    title = "Non-College and Hospitalization",
    x = "Non-College", 
    y = "Covid Hospitalization Rate per 100,000"
  ) +
  theme(legend.position = "right")
```



Insurance (Public and Private Insurance Included) and Vaccination Relationship 
```{r}
insurance_vax_rel <- 
  borough_info %>% 
  ggplot(aes(x = perc_insured,
             y = covid_vax_rate)) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(
    method = "loess", 
    color = 'red',
    se = FALSE) +
  labs(
    title = "Insurance VS.COVID Vaccination Rate", 
    subtitle = "By NYC Borough", 
    x = "Percentage Insured",
    y = "COVID Vaccination Rate") +
  theme(legend.position = "right")
```


English Speaker and Vaccination Relationship 
```{r}
english_vax_rel <- 
  borough_info %>% 
  ggplot(aes(x = perc_english,
             y = covid_vax_rate)) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(
    method = "lm", 
    color = 'red',
    se = FALSE) +
  labs(
    title = "English Speaking", 
    subtitle = "By NYC Borough", 
    x = "Percentage Speaking English",
    y = "COVID Vaccination Rate") +
  theme(legend.position = "none")
```



Non-English Speaker and Vaccination Relationship 
```{r}
nonenglish_vax_rel <- 
  borough_info %>% 
  ggplot(aes(x = perc_non_english,
             y = covid_vax_rate)) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(
    method = "lm", 
    color = 'red',
    se = FALSE) +
  labs(
    title = "Non - English Speaking", 
    subtitle = "By NYC Borough", 
    x = "Percentage Speaking English",
    y = "COVID Vaccination Rate") +
  theme(legend.position = "right")
```




Compare the (language spoken) plots side by side 
```{r}
english_vax_rel + nonenglish_vax_rel
```




USA Born and Vaccination Relationship 
```{r}
usaborn_vax_rel <- 
  borough_info %>% 
  ggplot(aes(x = perc_us_born,
             y = covid_vax_rate)) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE) + 
  labs(
    title = "US Born", 
    subtitle = "By NYC Borough", 
    x = "Percentage U.S. Born",
    y = "COVID Vaccination Rate") +
  theme(legend.position = "none")
```



Foreign and Vaccination Relationship 
```{r}
foreignborn_vax_rel <- 
  borough_info %>% 
  ggplot(aes(x = perc_foreign_born,
             y = covid_vax_rate)) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE) +
  labs(
    title = "Foreign Born", 
    subtitle = "By NYC Borough", 
    x = "Percentage Foreign Born",
    y = "COVID Vaccination Rate") +
  theme(legend.position = "right")
```




Compare the U.S. and Foreign born plots side by side 
```{r}
usaborn_vax_rel + foreignborn_vax_rel
```



Internet and Vaccination Relationship
```{r}
internet_vax_rel <- 
  borough_info %>% 
  ggplot(aes(x = perc_broadband,
             y = covid_vax_rate)) + 
  geom_point(aes(color  = Borough)) + 
  geom_smooth(
    method = "lm", 
    color = 'red',
    se = FALSE) +
  labs(
    title = "Internet and Vaccination Rate", 
    subtitle = "By NYC Borough", 
    x = "Internet Percentages for Each Puma in their associated Borough",
    y = "COVID Vaccination Rate") +
  theme(legend.position = "right")
```


Internet and Hospitalization Relationship
```{r}
internet_hosp_rel <- 
  borough_info %>% 
  ggplot(aes(x = perc_broadband,
             y = covid_hosp_rate)) + 
  geom_point(aes(color  = Borough)) + 
  geom_smooth(
    method = "lm", 
    color = 'red',
    se = FALSE) +
  labs(
    title = "Internet and Hospitalization Rate", 
    subtitle = "By NYC Borough", 
    x = "Internet Percentages for Each Puma in their associated Borough",
    y = "COVID Hospitalization Rate") +
  theme(legend.position = "right")
```



Race (White) vs Hospitalization Rate 
```{r}
borough_info %>% 
  ggplot(aes(
    x = perc_white,
    y = covid_hosp_rate
  )) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(method = "loess", 
    color = 'red',
    se = FALSE) +  
  labs(
    title = "Race (White) and Hospitalization Rate", 
    subtitle = "By NYC Borough", 
    x = "Percentage of White in that PUMA for that Borough",
    y = "COVID Hospitalization Rate")
```


Race (White) vs Vaccination Rate
```{r}
borough_info %>% 
  ggplot(aes(
    x = perc_white,
    y = covid_vax_rate
  )) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(method = "loess", 
    color = 'red',
    se = FALSE) +  
  labs(
    title = "Race (White) and Hospitalization Rate", 
    subtitle = "By NYC Borough", 
    x = "Percentage of White in that PUMA for that Borough",
    y = "COVID Vaccination Rate")
```



Race (White) vs Death Rate 
```{r}
borough_info %>% 
  ggplot(aes(
    x = perc_white,
    y = covid_death_rate
  )) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(method = "loess", 
    color = 'red',
    se = FALSE) +  
  labs(
    title = "Race (White) and Hospitalization Rate", 
    subtitle = "By NYC Borough", 
    x = "Percentage of White in that PUMA for that Borough",
    y = "COVID Death Rate")
```




Ranking -  Vaccination Rate in Puma (colored by Borough)
```{r}
rank_vax_puma <- 
  borough_info %>% 
  arrange(desc(covid_vax_rate)) %>%
  ggplot(aes(x = fct_reorder(puma, 
                             desc(covid_vax_rate)), 
             y = covid_vax_rate)) + 
  geom_bar(stat = "identity", 
           aes(fill = Borough)) + 
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, hjust = 0.5)) + 
  labs(
    title = "Ranking Vaccination Rate in each Puma", 
    x = "PUMA",
    y = "COVID Vaccination rate"
  ) + 
  theme(legend.position = "bottom")
```




Ranking -  Death Rate in Puma (colored by Borough)
```{r}
rank_death_puma <- 
  borough_info %>% 
  arrange(desc(covid_death_rate)) %>%
  ggplot(aes(x = fct_reorder(puma, 
                             desc(covid_death_rate)), 
             y = covid_death_rate)) + 
  geom_bar(stat = "identity", 
           aes(fill = Borough)) + 
  theme(axis.text.x = 
          element_text(angle = 70, vjust = 0.5, hjust = 0.5)) + 
  labs(
    title = "Ranking Death Rate in each Puma", 
    x = "PUMA",
    y = "COVID Death Rate"
  ) + 
  theme(legend.position = "bottom")
```




Ranking - Hospitalization Rate in Puma (colored by Borough)
```{r}
rank_hospr_puma <- 
  borough_info %>% 
  arrange(desc(covid_hosp_rate)) %>%
  ggplot(aes(x = fct_reorder(puma, 
                             desc(covid_hosp_rate)), 
             y = covid_hosp_rate)) + 
  geom_bar(stat = "identity", 
           aes(fill = Borough)) + 
  theme(axis.text.x = 
          element_text(angle = 70, vjust = 0.5, hjust = 0.5)) + 
  labs(
    title = "Ranking Hospitalization Rate in each Puma", 
    x = "PUMA",
    y = "COVID Hospitalization Rate"
  ) + 
  theme(legend.position = "bottom") + 
  facet_grid(Borough~puma)
```




```{r}
borough_info %>% 
  ggplot(aes(
    x = perc_male,
    y = covid_hosp_rate
  )) + 
  geom_point() + 
  geom_smooth(method = lm, 
              formula = y ~ splines::bs(x, 3), 
              se = FALSE) + 
   labs(
    title = "Relationship between Male and Hospitalization Rate", 
    x = "Percentage of Male in Each Puma",
    y = "COVID Hospitalization Rate") 
```



# Meeting Discussion - Creating Plots - *(11/30/2021)* - Verifying the Middle Heat Map in Zak's Document 

*Hospitalization Rate*

WHite and Hospitalization Rate 
```{r}
borough_info %>% 
  ggplot(aes(
    x = perc_white,
    y = covid_hosp_rate
  )) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(method = "loess", 
    color = 'red',
    se = FALSE) +  
  labs(
    title = "Race (White) and Hospitalization Rate", 
    subtitle = "By NYC Borough", 
    x = "Percentage of White in that PUMA for that Borough",
    y = "COVID Hospitalization Rate")
```


Public Transport and Hospitalization Rate 
```{r}
borough_info %>% 
  ggplot(aes(
    x = perc_public_transport,
    y = covid_hosp_rate
  )) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(method = "loess", 
    color = 'red',
    se = FALSE) +  
  labs(
    title = "Public Transportation and Hospitalization Rate", 
    subtitle = "By NYC Borough", 
    x = "Percentage of Public Transportation usage in that PUMA for that Borough",
    y = "COVID Hospitalization Rate")
```



Insured and Hospitalization Rate 
```{r}
borough_info %>% 
  ggplot(aes(
    x = perc_insured,
    y = covid_hosp_rate
  )) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(method = "loess", 
    color = 'red',
    se = FALSE) +  
  labs(
    title = "Insured and Hospitalization Rate", 
    subtitle = "By NYC Borough", 
    x = "Percentage of Insured in that PUMA for that Borough",
    y = "COVID Hospitalization Rate")
```



English and Hospitalization Rate 
```{r}
english_vax_rel <- 
  borough_info %>% 
  ggplot(aes(x = perc_english,
             y = covid_hosp_rate)) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(
    method = "lm", 
    color = 'red',
    se = FALSE) +
  labs(
    title = "English Speaking and Hospitalization Rate", 
    subtitle = "By NYC Borough", 
    x = "Percentage Speaking English in that puma for that Borough",
    y = "COVID Hospitalization Rate") +
  theme(legend.position = "none")
```


College and Hospitalization Rate 
```{r}
college_hosp_rel <- 
  borough_info %>% 
  ggplot(aes(x = perc_college,
             y = covid_hosp_rate)) + 
  geom_point(aes(color = Borough)) + 
  geom_smooth(method = lm, formula = y ~ splines::bs(x, 3), se = FALSE) + 
  labs(
    title = "College and Hospitalization",
    x = "College", 
    y = "Covid Hospitalization Rate per 100,000"
  ) +
  theme(legend.position = "right")
```


Internet and Hospitalization Relationship
```{r}
internet_hosp_rel <- 
  borough_info %>% 
  ggplot(aes(x = perc_broadband,
             y = covid_hosp_rate)) + 
  geom_point(aes(color  = Borough)) + 
  geom_smooth(
    method = "lm", 
    color = 'red',
    se = FALSE) +
  labs(
    title = "Internet and Hospitalization Rate", 
    subtitle = "By NYC Borough", 
    x = "Internet Percentages for Each Puma in their associated Borough",
    y = "COVID Hospitalization Rate") +
  theme(legend.position = "right")
```


Median Personal Income and Covid Hospitalization Rate 
```{r}
borough_info %>% 
  ggplot(aes(x = median_personal_income,
             y = covid_hosp_rate)) + 
  geom_point(aes(color  = Borough)) + 
  geom_smooth(
    method = "lm", 
    color = 'red',
    se = FALSE) +
  labs(
    title = "Median Personal Income and Hospitalization Rate", 
    subtitle = "By NYC Borough", 
    x = "Median Personal Income",
    y = "COVID Hospitalization Rate") +
  theme(legend.position = "right")
```


Median Household Income and Covid Hospitalization Rate
```{r}
borough_info %>% 
  ggplot(aes(x = median_household_income,
             y = covid_hosp_rate)) + 
  geom_point(aes(color  = Borough)) + 
  geom_smooth(
    method = "lm", 
    color = 'red',
    se = FALSE) +
  labs(
    title = "Median Household Income and Hospitalization Rate", 
    subtitle = "By NYC Borough", 
    x = "Median Household Income",
    y = "COVID Hospitalization Rate") +
  theme(legend.position = "right")
```


*Death Rate* 

Median Household income and Covid Death Rate 
```{r}
borough_info %>% 
  ggplot(aes(x = median_household_income,
             y = covid_death_rate)) + 
  geom_point(aes(color  = Borough)) + 
  geom_smooth(
    method = "lm", 
    color = 'red',
    se = FALSE) +
  labs(
    title = "Median Household Income and Death Rate", 
    subtitle = "By NYC Borough", 
    x = "Median Household Income in that Puma for that Borough",
    y = "COVID Death Rate") +
  theme(legend.position = "right")
```


Median Personal Income and Covid Death Rate 
```{r}
borough_info %>% 
  ggplot(aes(x = median_personal_income,
             y = covid_death_rate)) + 
  geom_point(aes(color  = Borough)) + 
  geom_smooth(
    method = "lm", 
    color = 'red',
    se = FALSE) +
  labs(
    title = "Median Personal Income and Death Rate", 
    subtitle = "By NYC Borough", 
    x = "Median Personal Income in that Puma for that Borough",
    y = "COVID Death Rate") +
  theme(legend.position = "right")
```



