---
title: "P8105: Data Science I"
author: "Exploration of Datasets<br>Jimmy Kelliher (UNI: jmk2303)"
output:
  github_document:
    toc: TRUE
---

<!-------------------------------------------------------------------------------
Preamble
-------------------------------------------------------------------------------->

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# load necessary packages
library(p8105.datasets)
library(tidyverse)

# set knitr defaults
knitr::opts_chunk$set(
    echo      = TRUE
  , message   = FALSE
  , fig.width = 6
  , fig.asp   = .6
  , out.width = "90%"
)

# set theme defaults
theme_set(
  theme_bw() +
  theme(
    legend.position = "bottom"
    , plot.title    = element_text(hjust = 0.5)
    , plot.subtitle = element_text(hjust = 0.5)    
    , plot.caption  = element_text(hjust = 0.0)
  )
)

# set color scale defaults
options(
    ggplot2.continuous.colour = "viridis"
  , ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```

<!-------------------------------------------------------------------------------
Census Data
-------------------------------------------------------------------------------->

# NYC Census Data

We pull demographic data from the 2019 5-year ACS for NYC. Below, we further filter by taking a random sample of 1,000 observations.

```{r}
read_csv("nyc_census_2019_sample.csv") %>%
  janitor::clean_names() %>%
  head(10) %>%
  knitr::kable()
```

# Health Outcomes

We pull hospitalizations and deaths per 100,000 residents in a given ZCTA from NYC DOHMH.

```{r}
read_csv("nyc_health_outcomes_2020.csv") %>%
  janitor::clean_names() %>%
  head(10) %>%
  knitr::kable()
```

# Crosswalk

Because the census data is observed at the PUMA geography, and because the health data is observed at the ZCTA geography, we pull a crosswalk between these two geographies. Note that neither geography is a strict subset of the other.

```{r}
read_csv("nyc_puma_zcta_crosswalk.csv") %>%
  janitor::clean_names() %>%
  head(10) %>%
  knitr::kable()
```