---
title: "P8105: Data Science I"
author: "Analysis Playground<br>Jimmy Kelliher (UNI: jmk2303)"
output:
  github_document:
    toc: TRUE
---

<!------------------------------------------------------------------------------
Preamble
------------------------------------------------------------------------------->

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# load necessary packages
library(tidyverse)

# set knitr defaults
knitr::opts_chunk$set(
    echo      = TRUE
  , message   = FALSE
  , fig.width = 6
  , fig.asp   = .6
  , out.width = "90%"
)

# set theme defaults
theme_set(
  theme_bw() +
  theme(
    legend.position = "bottom"
    , plot.title    = element_text(hjust = 0.5)
    , plot.subtitle = element_text(hjust = 0.5)    
    , plot.caption  = element_text(hjust = 0.0)
  )
)

# set color scale defaults
options(
    ggplot2.continuous.colour = "viridis"
  , ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```

<!------------------------------------------------------------------------------
Overview
------------------------------------------------------------------------------->

This file provides a template by which we create our final dataset. Essentially, because of the size of our census data, we cannot store the final dataset locally. Instead, we create a dataset of health outcomes at the PUMA level (see `data_documentation/merging_data_sets.Rmd`), and we merge this with our unzipped census data, which exists at the individual level and indicates the PUMA in which each individual resides.

# Unzipping the Census Data

```{r}
jimzip <- function(csv_file, path) {
  # create full path to csv file
  full_csv <- paste0(path, "/", csv_file)
  # append ".zip" to csv file
  zip_file <- paste0(full_csv, ".zip")
  # unzip file
  unzip(zip_file)
  # read csv
  data_extract <- read_csv(csv_file)
  # be sure to remove file once unzipped (it will live in working directory)
  on.exit(file.remove(csv_file))
  # output data
  data_extract
}

census_data <- jimzip("census_filtered.csv", "./data")
head(census_data) %>%  knitr::kable()
```

# Merging the Outcome Data

```{r}
health_data <-
  read_csv("./data/outcome_puma.csv") %>%
  rename(puma = puma10)

merged_data <- merge(census_data, health_data, by = "puma")
rm(census_data)

head(merged_data) %>% knitr::kable()
```

# Illustrating with an Example Analysis

```{r}
example <-
  merged_data %>%
  # restrict to white New Yorkers
  filter(race %in% c(1)) %>%
  # group by puma, race, and outcomme of interest
  group_by(puma, race, puma_hosp_rate) %>%
  # summarize predictors of interest by taking weighted averages
  summarize(
      age    = weighted.mean(age,     perwt, na.rm = TRUE)
    , educ   = weighted.mean(educ,    perwt, na.rm = TRUE)
    , inctot = weighted.mean(inctot,  perwt, na.rm = TRUE)
  )

# run a linear regression on stratified sample
summary(lm(puma_hosp_rate ~ age + educ + inctot, data = example))
``` 