---
title: "plot_ly first draft"
output: github_document
---


```{r message = FALSE, echo = FALSE}
library(tidyverse)
library(plotly)
library(ggplot2)
```



*CREATING THE PLOTLY WITH THE EXPLORATORY ANALYSIS FILE, NOT FINAL, STILL NEED TO ADD AND EDIT* 


**General Trend for each PUMA in each borough: Vaccination Rate, Death Rate, Hospitalization Rate for each PUMA in each Borough**

# Hospitalization rates across PUMAs, colored by borough
```{r include = FALSE}
puma_hosp <-
  puma_level_data %>% 
  mutate(puma = fct_reorder(puma, covid_hosp_rate)) %>% 
  mutate(text_label = str_c("PUMA: ", puma, 
                            "\n% Hospitalization: ", covid_hosp_rate)) %>%
  plot_ly(y = ~covid_hosp_rate, 
          x = ~puma, 
          color = ~borough, 
          type = "scatter",
          mode = "markers", 
          colors = "viridis",
          text = ~ text_label) %>% 
  config(scrollZoom = FALSE,
         modeBarButtonsToAdd = list('drawline', 
                                 'drawopenpath', 
                                 'drawclosedpath', 
                                 'drawcircle', 
                                 'drawrect', 
                                 'eraseshape'),
         displayModeBar = TRUE) %>%
  layout(plot_bgcolor = '#e5ecf6',
          title = "Hospitalization rates across PUMAs", 
          xaxis = list(
            title = "PUMA", 
            tickangle = 60,
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
          yaxis = list(
            title = "% Hospitalized", 
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
         legend = list(title = list(text = 'Borough')))
```


# Death rates across PUMAs, colored by borough
```{r include = FALSE}
puma_death <-
  puma_level_data %>% 
  mutate(puma = fct_reorder(puma, covid_death_rate)) %>% 
  mutate(text_label = str_c("PUMA: ", puma, 
                            "\n% Death: ", covid_death_rate)) %>%
  plot_ly(y = ~covid_death_rate, 
          x = ~puma, 
          color = ~borough, 
          type = "scatter",
          mode = "markers", 
          colors = "viridis",
          text = ~ text_label) %>% 
  config(scrollZoom = FALSE,
         modeBarButtonsToAdd = list('drawline', 
                                 'drawopenpath', 
                                 'drawclosedpath', 
                                 'drawcircle', 
                                 'drawrect', 
                                 'eraseshape'),
         displayModeBar = TRUE) %>%
  layout(plot_bgcolor = '#e5ecf6',
          title = "Death rates across PUMAs", 
          xaxis = list(
            title = "PUMA", 
            tickangle = 60,
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
          yaxis = list(
            title = "% Death", 
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
         legend = list(title = list(text = 'Borough')))
```


# Vaccination rates across PUMAs, colored by borough
```{r include = FALSE}
puma_vax <-
  puma_level_data %>% 
  mutate(puma = fct_reorder(puma, covid_vax_rate)) %>% 
  mutate(text_label = str_c("PUMA: ", puma, 
                            "\n% Vaccinated: ", covid_vax_rate)) %>%
  plot_ly(y = ~covid_vax_rate, 
          x = ~puma, 
          color = ~borough, 
          type = "scatter",
          mode = "markers", 
          colors = "viridis",
          text = ~ text_label) %>% 
  config(scrollZoom = FALSE,
         modeBarButtonsToAdd = list('drawline', 
                                 'drawopenpath', 
                                 'drawclosedpath', 
                                 'drawcircle', 
                                 'drawrect', 
                                 'eraseshape'),
         displayModeBar = TRUE) %>%
  layout(plot_bgcolor = '#e5ecf6',
          title = "Vaccination rates across PUMAs", 
          xaxis = list(
            title = "PUMA", 
            tickangle = 60,
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
          yaxis = list(
            title = "% Death", 
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
         legend = list(title = list(text = 'Borough')))
```

**Which PUMAs have the highest Hospitalization Rate, Death Rate, Vaccination Rate**

# Puma with the highest hospitalization rate 
```{r include = FALSE}
puma_level_data %>% 
  arrange(desc(covid_hosp_rate)) %>% 
  mutate(rank 
         = rank(desc(covid_hosp_rate))) %>% 
  filter(rank < 11) %>% 
  mutate(puma = 
           fct_reorder(puma, desc(covid_hosp_rate))) %>%
  mutate(text_label = str_c("PUMA: ", puma, 
                            "\n% Hospitalization: ", covid_hosp_rate)) %>%
  plot_ly(x = ~puma,
          y = ~covid_hosp_rate, 
          color = ~borough, 
          type = "bar",
          text = ~ text_label, 
          colors = "viridis") %>% 
  layout(plot_bgcolor = '#e5ecf6',
          title = "Top 10 Puma with the highest Hospitalization Rate", 
          xaxis = list(
            title = "PUMA", 
            tickangle = 60,
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
          yaxis = list(
            title = "% Death", 
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
         legend = list(title = list(text = 'Borough')))
```




# Puma with the lowest Hospitalization Rate
```{r}
PUMA_hosp_lowest10 = 
  puma_level_data %>% 
  arrange(covid_hosp_rate) %>% 
  mutate(
    rank = rank(desc(covid_hosp_rate))
  ) %>% 
  filter(rank < 11) %>% 
  mutate(puma = fct_reorder(puma, covid_hosp_rate)) %>%
  mutate(text_label = str_c("PUMA: ", puma, 
                            "\n% Hospitalized: ", covid_hosp_rate)) %>%
  plot_ly(x = ~puma,
          y = ~covid_hosp_rate, 
          color = ~borough, 
          type = "bar",
          text = ~ text_label, 
          colors = "viridis") %>% 
  layout(plot_bgcolor = '#e5ecf6',
          title = "Top 10 Puma with the Lowest Hospitalization Rate", 
          xaxis = list(
            title = "PUMA", 
            tickangle = 60,
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
          yaxis = list(
            title = "% Death", 
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
         legend = list(title = list(text = 'Borough')))
```





# Puma with the highest Death rate 
```{r include = FALSE}
puma_level_data %>% 
  arrange(desc(covid_death_rate)) %>% 
  mutate(rank 
         = rank(desc(covid_death_rate))) %>% 
  filter(rank < 11) %>% 
  mutate(puma = 
           fct_reorder(puma, desc(covid_death_rate))) %>%
  mutate(text_label = str_c("PUMA: ", puma, 
                            "\n% Death: ", covid_death_rate)) %>%
  plot_ly(x = ~puma,
          y = ~covid_death_rate, 
          color = ~borough, 
          type = "bar",
          text = ~ text_label, 
          colors = "viridis") %>% 
  layout(plot_bgcolor = '#e5ecf6',
          title = "Top 10 Puma with the highest Death Rate", 
          xaxis = list(
            title = "PUMA", 
            tickangle = 60,
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
          yaxis = list(
            title = "% Death", 
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
         legend = list(title = list(text = 'Borough')))
```



# Puma with the lowest Death Rate
```{r}
PUMA_death_lowest10 = 
  puma_level_data %>% 
  arrange(covid_death_rate) %>% 
  mutate(
    rank = rank(desc(covid_death_rate))
  ) %>% 
  filter(rank < 11) %>% 
  mutate(puma = fct_reorder(puma, covid_death_rate)) %>%
  mutate(text_label = str_c("PUMA: ", puma, 
                            "\n% Death: ", covid_death_rate)) %>%
  plot_ly(x = ~puma,
          y = ~covid_death_rate, 
          color = ~borough, 
          type = "bar",
          text = ~ text_label, 
          colors = "viridis") %>% 
  layout(plot_bgcolor = '#e5ecf6',
          title = "Top 10 Puma with the lowest Death Rate", 
          xaxis = list(
            title = "PUMA", 
            tickangle = 60,
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
          yaxis = list(
            title = "% Death", 
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
         legend = list(title = list(text = 'Borough')))
```





# Puma with the highest Vaccination rate 
```{r include = FALSE}
puma_level_data %>% 
  arrange(desc(covid_vax_rate)) %>% 
  mutate(rank 
         = rank(desc(covid_vax_rate))) %>% 
  filter(rank < 11) %>% 
  mutate(puma = 
           fct_reorder(puma, desc(covid_vax_rate))) %>%
  mutate(text_label = str_c("PUMA: ", puma, 
                            "\n% Vaccinated: ", covid_vax_rate)) %>%
  plot_ly(x = ~puma,
          y = ~covid_vax_rate, 
          color = ~borough, 
          type = "bar",
          text = ~ text_label, 
          colors = "viridis") %>% 
  layout(plot_bgcolor = '#e5ecf6',
          title = "Top 10 Puma with the highest Vaccination Rate", 
          xaxis = list(
            title = "PUMA", 
            tickangle = 60,
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
          yaxis = list(
            title = "% Death", 
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
         legend = list(title = list(text = 'Borough')))
```




# PUMA with the lowest vaccination rate
```{r}
PUMA_vax_lowest10 =
  puma_level_data %>% 
  arrange(covid_vax_rate) %>% 
  mutate(
    rank = rank(covid_vax_rate)
  ) %>% 
  filter(rank < 11) %>% 
  mutate(puma = fct_reorder(puma, covid_vax_rate)) %>% 
  mutate(text_label = str_c("PUMA: ", puma, 
                            "\n% Vaccinated: ", covid_vax_rate)) %>%
  plot_ly(x = ~puma,
          y = ~covid_vax_rate, 
          color = ~borough, 
          type = "bar",
          text = ~text_label, 
          colors = "viridis") %>% 
  layout(plot_bgcolor = '#e5ecf6',
          title = "Top 10 Puma with the lowest Vaccination Rate", 
          xaxis = list(
            title = "PUMA", 
            tickangle = 60,
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
          yaxis = list(
            title = "% Death", 
            zerolinecolor = '#ffff',
            zerolinewidth = 2,
            gridcolor = 'ffff'),
         legend = list(title = list(text = 'Borough')))
```





# Welfare vs Vax
```{r}
vax_welfare = 
  puma_level_data %>% 
  ggplot(
    aes(
      x = perc_welfare,
      y = covid_vax_rate
    )
  ) + 
  geom_point(aes(color = borough)) + 
  geom_smooth(se = FALSE, method = "lm", color = "red", aes(weight = total_people)) + 
  labs(
    x = "% on Welfare",
    y = "% Vaccinated"
  ) + 
  xlim(0, 100)

vax_welfare_plotly =
  ggplotly(vax_welfare)
```



# Vax vs Foodstamps 
```{r}
vax_foodstamps = 
  puma_level_data %>% 
  ggplot(
    aes(
      x = perc_foodstamps,
      y = covid_vax_rate
    )
  ) + 
  geom_point(aes(color = borough)) + 
  geom_smooth(se = FALSE, method = "lm", color = "red", aes(weight = total_people)) + 
  labs(
    x = "% on Food Stamps",
    y = "% Vaccinated"
  ) 

vax_foodstmps_plotly = 
  ggplotly(vax_foodstamps)
```



# Vax vs Income 
```{r}
vax_income = 
  puma_level_data %>% 
  ggplot(
    aes(
      x = median_personal_income,
      y = covid_vax_rate
    )
  ) + 
  geom_point(aes(color = borough)) + 
  geom_smooth(se = FALSE, method = "lm", color = "red", aes(weight = total_people)) + 
  labs(
    x = "Median Personal Income ($)",
    y = "% Vaccinated"
  ) 

vax_income_plotly = 
  ggplotly(vax_income)
```



# Vax and Unemployment 
```{r}
vax_unemployment = 
  puma_level_data %>% 
  ggplot(
    aes(
      x = perc_unemployed,
      y = covid_vax_rate
    )
  ) + 
  geom_point(aes(color = borough)) + 
  geom_smooth(se = FALSE, method = "lm", color = "red", aes(weight = total_people)) + 
  labs(
    x = "% Unemployed",
    y = "% Vaccinated"
  ) 

vax_unemploy_plotly = ggplotly(vax_unemployment)
```

