---
title: "P8105: Data Science I"
author: "Analysis Playground<br>Zachary Katz (UNI: zak2132)"
output:
  github_document:
    toc: TRUE
---

<!------------------------------------------------------------------------------
Preamble
------------------------------------------------------------------------------->

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# load necessary packages
library(tidyverse)

# set knitr defaults
knitr::opts_chunk$set(
    echo      = TRUE
  , message   = FALSE
  , fig.width = 6
  , fig.asp   = .6
  , out.width = "90%"
)

# set theme defaults
theme_set(
  theme_bw() +
  theme(
    legend.position = "bottom"
    , plot.title    = element_text(hjust = 0.5)
    , plot.subtitle = element_text(hjust = 0.5)    
    , plot.caption  = element_text(hjust = 0.0)
  )
)

# set color scale defaults
options(
    ggplot2.continuous.colour = "viridis"
  , ggplot2.continuous.fill   = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete   = scale_fill_viridis_d
```

<!------------------------------------------------------------------------------
Overview
------------------------------------------------------------------------------->

This file provides a template by which we create our final dataset. Essentially, because of the size of our census data, we cannot store the final dataset locally. Instead, we create a dataset of health outcomes at the PUMA level (see `data_documentation/merging_data_sets.Rmd`), and we merge this with our unzipped census data, which exists at the individual level and indicates the PUMA in which each individual resides.

# Unzipping the Census Data

```{r}
jimzip <- function(csv_file, path) {
  # create full path to csv file
  full_csv <- paste0(path, "/", csv_file)
  # append ".zip" to csv file
  zip_file <- paste0(full_csv, ".zip")
  # unzip file
  unzip(zip_file)
  # read csv
  data_extract <- read_csv(csv_file)
  # be sure to remove file once unzipped (it will live in working directory)
  on.exit(file.remove(csv_file))
  # output data
  data_extract
}

census_data <- jimzip("census_filtered.csv", "./data")
head(census_data) %>%  knitr::kable()
```

# Merging the Outcome Data

```{r}
health_data <-
  read_csv("./data/outcome_puma.csv") %>%
  rename(puma = puma10)

merged_data <- merge(census_data, health_data, by = "puma")
rm(census_data)

head(merged_data) %>% knitr::kable()
```

# Cleaning the Outcome Data

```{r}
merged_data %>% head(10) %>% View()

cleaned_data = 
  merged_data %>% 
  # Eliminate duplicate rows
  distinct() %>% 
  # Rename variables
  rename(
    borough = countyfip,
    has_broadband = cihispeed,
    birthplace = bpl,
    years_in_usa = yrsusa1
  ) %>% 
  mutate(
    borough = recode(
      borough,
      "5" = "Bronx",
      "47" = "Brooklyn",
      "61" = "Manhattan",
      "81" = "Queens",
      "85" = "Staten Island"
    ),
    foodstmp = recode(
      foodstmp,
      "1" = "No",
      "2" = "Yes"
    ),
    has_broadband = case_when(
      has_broadband == "20" ~ "No",
      has_broadband != "20" ~ "Yes"
    ),
    sex = recode(
      sex,
      "1" = "Male",
      "2" = "Female"
    ),
    race = case_when(
      race == "1" ~ "White",
      race == "2" ~ "Black",
      race == "3" ~ "American Indian",
      race %in% c(4,5,6) ~ "Asian",
      race == "7" ~ "Other",
      race %in% c(8,9) ~ "2+ races"
    ),
    hispan = case_when(
      hispan == "0" ~ "No",
      hispan %in% c(1,2,3,4) ~ "Yes",
      hispan == "9" ~ "Not Reported"
    ),
    birthplace = case_when(
      birthplace %in% 1:120 ~"US",
      birthplace %in% 121:950 ~ "Non-US",
      birthplace == 999 ~"Unknown"
    ),
    citizen = case_when(
      citizen %in% c(1,2) ~ "Yes",
      citizen %in% 3:8 ~"No",
      citizen %in% c(0,9) ~ "Unknown"
    ),
    language = case_when(
      language == "1" ~ "English",
      language == "12" ~ "Spanish",
      language == "43" ~ "Chinese",
      language == "0" ~ "Unknown",
      language == "31" ~ "Hindi",
      !language %in% c(1,12,43,0,31) ~ "Other"
    )
  ) %>% 
  mutate(
    annual_rent = rent * 12
  ) %>% 
  mutate(
    health_insurance = case_when(
      hcovany == 1 ~ "None",
      hcovany == 2 && hcovpriv == 2 ~ "Private",
      hcovany == 2 && hcovpriv == 1 ~ "Public"
    )
  ) %>% 
  # Eliminate redundant variables
  select(-rent, -hispan, -ancestr1, -ancestr2, -hcovany, -hcovpriv, -hcovpub) %>% 
  head(10) %>% 
  View()


  # Optimize data types across variables
  mutate(
    puma = as.factor(puma),
    multyear = as.factor(multyear),
    foodstmp = as.factor(foodstmp),
    sex = as.factor(sex),
    race = as.factor(race)


%>% 
  mutate(
    borough = as.factor(borough),
    hispanic = as.factor(hispan),
    birthplace = as.factor(birthplace),
    citizen = 
  ) 

str(merged_data)

merged_data %>% group_by(language) %>% summarize(n = n()) %>% arrange(desc(n))
```

